---
title: "Reproducible Research - Week 2 Project"
author: "Aliza Shoop"
output: html_document
---
# Read Date & Load Libraries

```{r echo=TRUE}
setwd("~/dsCourses")
act <- read.csv("activity.csv")

library(ggplot2)
library(lubridate)
library(plyr)
```


## Copy data and clean up
```{r echo=TRUE}
act2 <- act
act2$date <- as.Date(act2$date)
#remove na's
act2 <-na.omit(act2)
```

# What is mean total number of steps taken per day?

## Mean and Median of steps
```{r echo=TRUE}
mean1 <- mean(act2$steps, na.rm = TRUE)
mean1
median1 <- median(act2$steps, na.rm = TRUE)
median1
```

## Create a histogram of steps
```{r echo=TRUE}
#assign a column with day number
act2$day<- day(act2$date)
#plot histogram
p1 <- ggplot(act2, aes(x=day, y=steps)) + 
     geom_histogram(stat="identity") + 
     ggtitle("Total number of steps taken per day") + 
     xlab("Day") + 
     ylab("Steps")

p1
```

# What is the average daily activity pattern?

## Get averages by interval

```{r echo=TRUE}
avg <- aggregate(act2$steps ~ act2$interval, FUN = mean)
names(avg) <- c("interval","steps")
```

## Plot a line chart

```{r echo=TRUE}
p2<- ggplot() + geom_line(aes(y=steps,x=interval),
                          data=avg)
p2+labs(title="Average Steps per Day",x="Interval", y="Steps")

```

## Which 5 minute interval has the highest average steps?

```{r echo=TRUE}
maxI<- avg[which.max(avg$steps),]

maxI
```

# Imputing missing values

```{r echo=TRUE}
#get total na's from original data
sum(is.na(act))
#copy data to new table
act3 <- act
#copy mean of steps to na's
for(i in 1:nrow(act3)){
    if(is.na(act3$steps[i])){
        act3$steps[i] <- mean(act3$steps,na.rm=TRUE)
    }
}
#check for na's in new data
sum(is.na(act3))
```
## Make a histogram of the new data

```{r echo=TRUE}
act3$day<- day(act3$date)

p3 <- ggplot(act3, aes(x=day, y=steps)) + 
     geom_histogram(stat="identity") + 
     ggtitle("Total number of steps taken per day") + 
     xlab("Day") + 
     ylab("Steps")

p3
```

## Calculate mean and median for new dataset
```{r echo=TRUE}
mean2 <-mean(act3$steps)
median2 <-median(act3$steps)
```
*Mean of original data `r mean1` vs mean of new data `r mean2`*  
*Median of original data `r median1` vs median of new data `r median2`*

# Are there differences in activity patterns between weekdays and weekends?

```{r echo=TRUE}
#create a weekday column
act3$date <- as.Date(act3$date)
act3$weekday <- ifelse(weekdays(act3$date) == "Saturday" | weekdays(act3$date) == "Sunday", "weekend", "weekday")

#calculate by interval
new_avg <- aggregate(act3$steps, list(act3$interval, act3$weekday), FUN = mean)
names(new_avg) <- c("steps", "weekday", "interval")
```

## Plot the difference between weekdays and weekends
```{r echo=TRUE}
ggplot(new_avg, aes(x=steps, y=interval)) + 
  geom_line() + 
  xlab("Interval") + 
  ylab("Average Steps") + 
  facet_grid(.~weekday)
```